Ephemeral hierarchical data-management for neuroimaging applications

### 2023-06-28 paper/sea-neuro/paper-neuro.tex
Minor updates to text

\documentclass[fleqn,10pt]{wlscirep}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{lineno}
\usepackage[binary-units]{siunitx}
\usepackage{multirow}
\usepackage{graphicx}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{xspace}

\newcommand{\TG}[1]{\color{red}\textsc{From Tristan}: #1\xspace\color{black}}

@@

	Neuroimaging open-data initiatives have led to increased availability of large
  scientific datasets. While these datasets are shifting the processing
  bottleneck from compute-intensive to data-intensive, current standardized
  analysis tools have yet to adopt strategies that mitigate the costs
  associated with large data transfers. A major challenge in adapting neuroimaging
  applications for data-intensive processing is that they must be entirely
  rewritten.

  To facilitate data management for standardized neuroimaging tools, we
  developed Sea, a library that intercepts and redirects application read and
  write calls to minimize data transfer time. In this paper, we investigate the
  performance of Sea on three preprocessing pipelines implemented using standard toolboxes (FSL,
  SPM and AFNI), using three neuroimaging datasets of different sizes
  (OpenNeuro's ds001545, PREVENT-AD and the HCP dataset) on two high-performance
  computing clusters. Our results demonstrate that given the features of the
  applications and datasets, Sea provides large speedups (up to 32$\times$) when
  the shared file system's (e.g. Lustre) performance is deteriorated. When the
  shared file system is not overburdened by other users, performance is
  unaffected by Sea, suggesting that Sea's overhead is minimal even in cases
  where its benefits are limited. Overall, Sea is beneficial, even when performance gain
  is minimal, as it can be used to limit files on PFS.
\end{abstract}
\begin{document}

@@

    Newer neuroimaging applications leveraging popular neuroimaging pipeline engines
    also do not benefit from processing-related data management.
    Although engines such as Nipype~\cite{nipype}
    % and Joblib~\cite{joblib} \TG{joblib is multithreading, dask or ray. I wouldn't call it an engine. You should use another example.}
    do not prohibit the use of data-management
    strategies, they do not facilitate the integration of these strategies into
    their resulting workflow. To give
    neuroimaging applications data management capabilities, the applications
    must interact with a file system or library that enable the strategies.

@@

    HPC systems rely on scalable network-based parallel file systems (e.g.,
    Lustre) for storage. While such file systems offer excellent performance,
    they are shared between all the users on the cluster, meaning that users
    with data-intensive applications can effectively deteriorate the performance
    of the shared file system for all users on the cluster. Solutions for
    improving shared file system performance include
    throttling~\cite{huang2020ooops} the data-intensive workloads or
    recommending the use of Burst Buffers~\cite{bb} (e.g., reserving a compute
    node for storage or leveraging local compute storage during processing). The
    latter, however, requires that the user manages their data to and from the
    Burst Buffer, unless a Burst Buffer file system, such as
    BurstFS~\cite{burstfs} or GekkoFS~\cite{gekkofs} is installed.

    In order for a file system to be mounted by an HPC cluster user, it must be
    loadable without administrative privileges. As the applications are
    typically made to interact with POSIX-based file systems, the new file
    system must also be compliant to the format. One method to ensure that these
    conditions are met is through the \texttt{LD\_PRELOAD} trick. This trick is
    used to intercept select library calls and redefine their behaviour. It has
    been used in many projects and to create lightweight userspace versions of
    file systems~\cite{xtreemfs,burstfs,gekkofs}.

@@

    \centering
    \includegraphics[width=\columnwidth]{figures/sea-diagram.pdf}%
\caption{The Sea data-management library overview}
\label{fig:seaneuro:diagram}
\end{figure*}
    Sea (Figure~\ref{fig:seaneuro:diagram}) is a data-management library that leverages
    the \texttt{LD\_PRELOAD}
    trick to intercept POSIX file system (more specifically, the GNU C library,
    glibc) calls on Linux systems. This enables Sea to redirect write calls
    aimed at slower storage devices to a faster devices whenever possible.
    Similarly, when intercepting read calls, Sea can choose to read from a
    faster device if a copy is available on that device. Sea decides which
    storage location it can write to based on the details provided in an
    initialization file called \texttt{sea.ini}. This file informs Sea of which
    locations it can use to read and write to, and well as their order of
    priority. It is important to note that Sea is not a filesystem, but rather a
    lightweight tool that redirects files to more optimal storage devices.

@@

    As neuroimaging pipeline results are typically required post-processing and
    HPC compute-local resources are only accessible during the reserved
    duration, Sea provides functionality to flush and evict data to persistent
    shared storage. This is accomplished via a separate thread (known as the
    ``flusher'') that moves data from the caches to long-term storage. A
    separate thread is used, in this case, to avoid interrupting ongoing
    processing with data management operations. Users must inform Sea of files
    that need to be persisted to storage within a file called
    \texttt{.sea\_flushlist}, and temporary files which can be removed from
    cache within a file called the \texttt{.sea\_evictlist}. Both these files
    can be populated using regular expressions to denote the paths to flush and
    evict. If a file occurs in both the \texttt{.sea\_flushlist} and
    \texttt{.sea\_evictlist}, Sea will interpret this as a \texttt{move}
    operation and simply move the file from local to persistent storage rather
    than copying it. Files that will be reused by the application should only be
    flushed rather than flushed and evicted, as files can benefit from speedups
    related to reading from fast rather than slow storage. Sea currently also
    provides a rudimentary prefetch thread that can move files located within
    Sea to the fastest available cache. To use Sea's prefetch
    capabilities, a file called \texttt{.sea\_prefetchlist} needs to be
    populated using regular expressions like the flushing and eviction files.

@@

\begin{subfigure}{0.5\textwidth}
    \centering
    \captionsetup{width=.85\linewidth}
    \includegraphics[width=\columnwidth]{figures/slashbin_nozoom.pdf}%
    \caption{Complete results}\label{fig:seaneuro:slashbinfull}
\end{subfigure}
\begin{subfigure}{0.5\textwidth}
    \centering
    \captionsetup{width=.85\linewidth}
    \includegraphics[width=\linewidth]{figures/slashbin_withzoom.pdf}
    \caption{Zoomed between 0 and 5000 seconds}\label{fig:seaneuro:slashbinzoom}
\end{subfigure}
\caption{Makespan comparison between Sea and Baseline on controlled dedicated cluster}
\label{fig:seaneuro:slashbin}
\end{figure*}

@@

In Figure~\ref{fig:seaneuro:slashbin}, we observe that although performance is
variable, speedups are obtained by using Sea when Lustre
performance has been degraded by busy writers. On average, a single SPM
process preprocessing a single HCP image produced the greatest speedup
(13$\times$). AFNI preprocessing a single fMRI image of the PREVENT-AD dataset
using a single process was the next fastest pipeline, with an average speedup of
(5$\times$). FSL Feat had speedups as well, with a maximum average speedup of
1.3$\times$ when preprocessing a single PREVENT-AD image. SPM consistently had
excellent speedups, which is likely due to a mix of prefetching the initial
input files and the I/O patterns of the application. While it was expected that
AFNI would benefit the most from Sea due to its
minimal compute time and the amount of data it generates, it did not benefit
from Sea to the extent that SPM did. As per Table~\ref{table:seaneuro:pipelines},
AFNI performs a very high number of glibc calls. While the overhead of an
individual call is likely very minimal, it is possible that the culmination of
all these calls results in a large enough bottleneck that affects both Sea and
Baseline executions an extensive amount. Moreover, the number of glibc calls
aimed at interacting with data on the Lustre storage are very few. AFNI still
obtained an average of 2$\times$ speedup across all experimental conditions.

FSL Feat, in contrast, appears to be the most compute-bound of the three
applications. Not only did it spend an extensive amount of time computing, the
amount of output data generated was least of the three. Due to the
compute-intensive nature of FSL Feat, it is expected that it resulted in the
least significant speedups.

Our results indicate that HCP, on average, obtained the greatest speedups with
Sea. Out of the three datasets, HCP has the largest images (see
Table~\ref{table:sea-neuro:data}). Larger individual images mean that they
occupy more page cache space and take longer to flush to Lustre. Unsurprisingly,
the next dataset with the largest images (\SI{282}{\mebi\byte} compressed for a
single image) results in the next largest speedups, followed by PREVENT-AD, the
dataset with the smallest functional images.

For both the pipelines and the datasets, we noticed that obtained speedups
decrease with increased parallelism. We did not modify the default parallelism
level of the applications in our experiments. As a result, each application
process is likely attempting to utilize all cores resulting in longer wait times
with the increase in threads. With this unintentional additional wait time that
is added to the applications due to compute contention, the benefits of using
Sea becomes more limited.

@@

Baseline performance without busy writers is comparable to that of Sea's as can be seen in Figure~\ref{fig:seaneuro:slashbin}. However, with
busy writers, Baseline performance may deviate dramatically from the identity
line. We observe that the more Baseline deviates from this identity line, the
greater the speedup.

We do occasionally observe slowdowns from using Sea. These slowdowns
may arise from the initial read of the data, as they appear to occur less
frequently with SPM, or due to increased CPU contention caused by Sea's rapid I/O.

\subsection{Overhead of Sea is minimal}

@@

To determine the overhead of Sea, we compared the average performance of Sea with
the performance of pipeline executing entirely within memory
(Figure~\ref{fig:seaneuro:tmpfs}). The results demonstrate that Sea and tmpfs
behave more or less identically. Given these results, we conclude that the
overhead of Sea is minimal and that the differences in makespan may be
due to Sea's initial reading of the data from Lustre.

@@

\begin{figure*}
\begin{subfigure}{0.5\textwidth}
    \centering
    \captionsetup{width=.85\linewidth}
    \includegraphics[width=\columnwidth]{figures/beluga_withflush_nozoom.pdf}%
    \caption{Complete results}\label{fig:seaneuro:belugafullwf}
\end{subfigure}
\begin{subfigure}{0.5\textwidth}
    \centering
    \captionsetup{width=.85\linewidth}
    \includegraphics[width=\linewidth]{figures/beluga_withflush_zoom.pdf}
    \caption{Zoomed between 0 and 2000 seconds}\label{fig:seaneuro:belugazoomwf}
\end{subfigure}
\caption{Makespan comparison between Sea and Baseline on the production cluster when flushing on all files is enabled}
\label{fig:seaneuro:beluga-wflush}
\end{figure*}

  Figures~\ref{fig:seaneuro:beluga-noflush} and \ref{fig:seaneuro:beluga-wflush}
  shows us that although performance in production clusters is highly variable,
  we can still obtain important speedups. It is very likely that when we
  executed our experiments without flushing
  (Figure~\ref{fig:seaneuro:beluga-noflush}), Lustre performance was not degraded, 
  resulting in Sea and Baseline performing
  quite similarly to each other.

  Our results with flushing enabled confirm this theory. When running with
  flushing enabled, which would have the additional overhead of having to
  ensure that all the data is copied to Lustre, we notice that occasionally, we
  obtain very large speedups with both SPM and AFNI, suggesting that Lustre
  performance was degraded. The
  maximum speedup observed when running on the production cluster with flushing enabled was
  11$\times$ with AFNI processing 1 HCP image. The maximum speedup for SPM was
  3.2$\times$, processing 8 ds001545 images.

@@

    Our results show that the speedups obtained from using Sea can be quite
    substantial (up to 32$\times$) depending on the degradation of the Lustre
    file system. The highest speed-ups observed on a production cluster were
    found to be 11$\times$.

@@

    Although Sea usage did result in performance degradation at times, the magnitude
    of the degradation is less than that of possible speedups
    obtained. Our controlled experiments demonstrate that using Sea with a
    degraded Lustre file system can result in average speedups of up to
    2.5$\times$. Such results indicate that the risk of a slowdown with Sea are
    less important that the potential speedups that can arise.

@@

    Sea really shines when it is executed on data-intensive pipelines. This
    means, the data is written at a rate which exceeds the rate at which the
    compute node's page cache can flush results to Lustre. In our experiments,
    the pipeline which consisted of both the shortest duration and largest
    output size was 16 AFNI processes preprocessing 16 HCP fMRI images. While the concurrent processes wrote, in total,
    approximately 100~GB of data, the controlled cluster would have had
    approximately 44~GB of dirty cache available per OST and a maximum amount of
    100~GB of page cache. Since all writes would not have occurred at the same
    time and computation time was extended as a result of CPU contention, even
    our controlled experiments do not showcase the benefits of Sea in a highly
    data-intensive scenario. However, HCP is one of the largest fMRI datasets,
    thus the only way to augment data intensiveness of the applications is
    through increased parallelism which is limited by the number of available
    cores. As a result, we can only demonstrate the speedups brought by Sea when
    Lustre performance has been degraded.

    Sea cannot speedup applications that are very compute-intensive, as is the case with FSL. This is because the bottleneck is the
    CPU time, which Sea does not address. Moreover, depending on the
    application's I/O patterns, it is possible that there will be sufficient
    page cache space and time to flush all the results asynchronously
    to Lustre. However, even with minimal speedups, Sea can still be useful in
    these cases to limit the number of new files created on Lustre, which can be
    particularly useful in instances where users have strict file limits.

@@

    Our results expose the complexities of performance predictions on HPC clusters. We present results on a dedicated cluster with which we have
    control over the global cluster usage and are able to demonstrate that
    significant speedups can be obtained when all the Lustre storage devices are
    busy writing other data concurrently. However, even these experiments are
    limited in demonstrating potential speedups with Sea as there are many
    controlled variables, some of which may vary from cluster to cluster. For
    example, the production cluster has a total of 977 nodes and only 38 disks associated with
    their Lustre scratch file system. We can imagine that if 900 of the nodes
    are busy writing to the scratch file system, we would observe a speedup
    larger than what has been reported on our dedicated cluster with only 8 compute
    nodes and 44 disks associated to Lustre. However, these are not the only
    factors that can affect reported results, others can include when the
    experiment were executed, what kind of applications were being executed
    alongside the experiment, scratch file system usage during experiments, and
    what were the application read and write patterns.

@@

    Sea provides prefetching, flushing and eviction mechanisms to further
    improve the speedups that can be obtained. Prefetching allows input data to
    be moved to and manipulated from memory. All of our experiments with SPM
    used prefetching as SPM updates the input data via the use of a memory map.
    Without prefetching, updates to the input files would have been performed
    directly on Lustre, thus exhibiting a less important speedup. It was
    favorable to proceed with prefetching, in this case, as the input files
    themselves were not large compared to available memory space. However, in
    cases with very large input files, it may be preferable to maintain the
    memory map to Lustre.

@@

    The ability to use glibc interception over more common alternative methods
    such as kernel-based or FUSE-based file systems was essential in ensuring
    ease-of-use and minimal performance overheads. As both our controlled and
    production environment experiments demonstrate, Sea's overhead is minimal
    and the magnitude of the speedups to be gained by the use of Sea is
    greater than the magnitude of the slowdowns.

@@

    From the perspective of a cluster administrator, Sea is considered useful to
    maintaining the global health of the system. By using available local
    resources, Sea is able to alleviate strain on globally shared cluster
    resources, thus improving the experience for all users. Sea is able to do
    this through the use of eviction which ensures that data that is not
    required, never makes it to Lustre.

@@

    To determine the performance gain Sea brings to neuroimaging analyses, we
    must evaluate the value of Sea on a variety of neuroimaging applications.
    For our analysis, we selected different fMRI preprocessing applications as
    fMRI processing has some of the most well-established tools for neuroimaging
    and some of the largest datasets. Of course, different modalities and tools
    may result in vastly different data access patterns and compute times, we
    use fMRI pre-processing as our baseline for the benefits that can be
    obtained with Sea.

@@

    Similarly to datasets, preprocessing pipelines can also vary in duration as
    a result of methodological differences. Thus, we preprocessed each dataset
    using three standard preprocessing pipelines available within the FSL~\cite{fsl},
    AFNI~\cite{cox1996afni}, and SPM~\cite{spm} toolboxes.
    Table~\ref{table:seaneuro:pipelines} provides a reference of the differences
    in computation and data-intensiveness of the different pipelines on a single
    subject for all three of the datasets.

    Each tool was set to only run the preprocessing pipeline. Scripts used to
    execute each pipeline can be found
    at~\url{github.com/valhayot/sea-slurm-scripts}.

    For the FSL preprocessing pipeline, we used FEAT configured with many of
    the default parameters. We set the following parameters: slice-timing
    correction (interleaved), intensity normalization and non-linear
    registration to standard space. FSL version 6.0.4 was used.

    %% Add SPM here
    To preprocess the datasets with SPM12, we reused the template described in
    \cite{haitas2021age}, with fieldmap correction removed. As the template was
    designed for a specific dataset, we updated the paths to the fMRI and
    anatomical images, as well as the dimensions, such as the number of volumes
    and slices and the TR and TA, for each dataset. Furthermore, we set the
    slice order to interleaved in all cases.

@@

    All the application Dockerfiles with Sea pre-installed are available
    at~\url{https://github.com/ValHayot/Sea/tree/master/containers/applications}.

@@

      \hline
    \end{tabular}}
    \footnotesize{$*$ Measurements taken on the dedicated cluster}
    \caption{Pipeline execution characteristics based on the processing of a single fMRI image using a single application process}\label{table:seaneuro:pipelines}
    \end{table*}

    \subsection{Controls}
    Our objective is to measure the speedup brought by Sea on real neuroimaging
    pipelines. Although Sea is agnostic to the internals of the pipelines and
    their output results, incorrect parameter selection and resulting outputs
    may result in I/O patterns that would not be observed had the pipeline been
    executed properly.

>>>>

\documentclass[fleqn,10pt]{wlscirep}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{lineno}
\usepackage[binary-units]{siunitx}
\usepackage{multirow}
\usepackage{graphicx}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{xspace}
\usepackage{todonotes}

\newcommand{\TG}[1]{\todo[color=green!60!black]{\textbf{\textsc{From Tristan:}} #1\xspace}}

@@

	Neuroimaging open-data initiatives have led to increased availability of large
  scientific datasets. While these datasets are shifting the processing
  scientific datasets. While these datasets are shifting the processing
  bottleneck from compute-intensive to data-intensive, current standardized
  analysis tools have yet to adopt strategies that mitigate the costs
  associated with large data transfers. A major challenge in adapting neuroimaging
  analysis tools have yet to adopt strategies that mitigate the costs
  associated with large data transfers. A major challenge in adapting neuroimaging
  applications for data-intensive processing is that they must be entirely
  rewritten. To facilitate data management for standardized neuroimaging tools, we
  developed Sea, a library that intercepts and redirects application read and
  write calls to minimize data transfer time. In this paper, we investigate the
  performance of Sea on three preprocessing pipelines implemented using standard toolboxes (FSL,
  SPM and AFNI), using three neuroimaging datasets of different sizes
  (OpenNeuro's ds001545, PREVENT-AD and the HCP dataset) on two high-performance
  computing clusters. Our results demonstrate that given the features of the
  applications and datasets, Sea provides large speedups (up to 32$\times$) when
  the shared file system's (e.g. Lustre) performance is deteriorated. When the
  shared file system is not overburdened by other users, performance is
  unaffected by Sea, suggesting that Sea's overhead is minimal even in cases
  where its benefits are limited. Overall, Sea is beneficial, even when performance gain
  is minimal, as it can be used to limit files on PFS.
\end{abstract}
\begin{document}

@@

    Newer neuroimaging applications leveraging popular neuroimaging pipeline engines
    also do not benefit from processing-related data management.
    Although engines such as Nipype~\cite{nipype}
    % and Joblib~\cite{joblib} \TG{joblib is multithreading, dask or ray. I wouldn't call it an engine. You should use another example.}
    do not prohibit the use of data-management
    strategies, they do not facilitate the integration of these strategies into
    their resulting workflow. To give
    Although engines such as Nipype~\cite{nipype}
    % and Joblib~\cite{joblib} \TG{joblib is multithreading, dask or ray. I wouldn't call it an engine. You should use another example.}
    do not prohibit the use of data-management
    strategies, they do not facilitate the integration of these strategies into
    their resulting workflow. To give
    neuroimaging applications data management capabilities, the applications
    must interact with a file system or library that enable the strategies.

@@

    HPC systems rely on scalable network-based parallel file systems (e.g.,
    HPC systems rely on scalable network-based parallel file systems (e.g.,
    Lustre) for storage. While such file systems offer excellent performance,
    they are shared between all the users on the cluster, meaning that users
    with data-intensive applications can effectively deteriorate the performance
    of the shared file system for all users on the cluster. Solutions for
    improving shared file system performance include
    throttling~\cite{huang2020ooops} the data-intensive workloads or
    recommending the use of Burst Buffers~\cite{bb} (e.g., reserving a compute
    recommending the use of Burst Buffers~\cite{bb} (e.g., reserving a compute
    node for storage or leveraging local compute storage during processing). The
    latter, however, requires that the user manages their data to and from the
    Burst Buffer, unless a Burst Buffer file system, such as
    BurstFS~\cite{burstfs} or GekkoFS~\cite{gekkofs} is installed.

    In order for a file system to be mounted by an HPC cluster user, it must be
    loadable without administrative privileges. As the applications are
    typically made to interact with POSIX-based file systems, the new file
    system must also be compliant to the format. One method to ensure that these
    conditions are met is through the \texttt{LD\_PRELOAD} trick. This trick is
    conditions are met is through the \texttt{LD\_PRELOAD} trick. This trick is
    used to intercept select library calls and redefine their behaviour. It has
    been used in many projects and to create lightweight userspace versions of
    file systems~\cite{xtreemfs,burstfs,gekkofs}.

@@

    \centering
    \includegraphics[width=\columnwidth]{figures/sea-diagram.pdf}%
\caption{The Sea data-management library overview}
\label{fig:seaneuro:diagram}
\end{figure*}
    Sea (Figure~\ref{fig:seaneuro:diagram}) is a data-management library that leverages
    the \texttt{LD\_PRELOAD}
    trick to intercept POSIX file system (more specifically, the GNU C library,
    glibc) calls on Linux systems. This enables Sea to redirect write calls
    aimed at slower storage devices to a faster devices whenever possible.
    Similarly, when intercepting read calls, Sea can choose to read from a
    faster device if a copy is available on that device. Sea decides which
    storage location it can write to based on the details provided in an
    storage location it can write to based on the details provided in an
    initialization file called \texttt{sea.ini}. This file informs Sea of which
    locations it can use to read and write to, and well as their order of
    priority. It is important to note that Sea is not a filesystem, but rather a
    lightweight tool that redirects files to more optimal storage devices.

@@

    As neuroimaging pipeline results typically require post-processing and
    HPC compute-local resources are only accessible during the reserved
    duration, Sea provides functionality to flush and evict data to persistent
    shared storage. To avoid interrupting ongoing
    processing with data management operations, this is accomplished via a separate thread (known as the
    ``flusher'') that moves data from the caches to long-term storage. Users must inform Sea of files
    that need to be persisted to storage within a file called
    \texttt{.sea\_flushlist}, and temporary files which can be removed from
    cache within a file called the \texttt{.sea\_evictlist}. Both these files
    can be populated using regular expressions to denote the paths to flush and
    evict. If a file occurs in both the \texttt{.sea\_flushlist} and
    \texttt{.sea\_evictlist}, Sea will interpret this as a \texttt{move}
    operation and simply move the file from local to persistent storage rather
    than copying it. Files that will be reused by the application should only be
    flushed rather than flushed and evicted, as files can benefit from speedups
    related to reading from fast rather than slow storage. Sea currently also
    provides a rudimentary prefetch thread that can move files located within
    Sea to the fastest available cache. To use Sea's prefetch
    capabilities, a file called \texttt{.sea\_prefetchlist} needs to be
    populated using regular expressions like the flushing and eviction files.

@@

\begin{subfigure}{0.5\textwidth}
    \centering
    \captionsetup{width=.85\linewidth}
    \includegraphics[width=\columnwidth]{figures/slashbin_nozoom.pdf}%
    \caption{Complete results}\label{fig:seaneuro:slashbinfull}
\end{subfigure}
\begin{subfigure}{0.5\textwidth}
    \centering
    \captionsetup{width=.85\linewidth}
    \includegraphics[width=\linewidth]{figures/slashbin_withzoom.pdf}
    \caption{Zoomed between 0 and 5000 seconds}\label{fig:seaneuro:slashbinzoom}
\end{subfigure}
\caption{Makespan comparison between Sea and Baseline on controlled dedicated cluster. Makespan
denotes the total time between execution launch and completion of the last computing task.}
\label{fig:seaneuro:slashbin}
\end{figure*}

@@

We benchmarked Sea by processing 3 functional MRI datasets of increasing sizes
(ds001545: 27~GB, PREVENT-AD: 283~GB, Human Connectome Project: 79~TB) with 3 pipelines 
from the most commonly-used neuroimaging toolboxes (AFNI, SPM, FSL), using a controlled HPC cluster 
dedicated to our experiments. We deployed ``busy writers'' in the cluster
that simulated the load from other user applications by writing to the shared file system at a controlled rate.

Speedups were obtained by using Sea when Lustre
performance had been degraded by busy writers (Figure~\ref{fig:seaneuro:slashbin}),
although performance was variable. On average, a single SPM pipeline
process preprocessing a single HCP image produced the greatest speedup
(13$\times$). The AFNI pipeline preprocessing a single fMRI image of the PREVENT-AD dataset
using a single process was the next fastest pipeline, with an average speedup of
(5$\times$). The FSL Feat pipeline had speedups as well, with a maximum average speedup of
1.3$\times$ when preprocessing a single PREVENT-AD image. The SPM pipeline consistently had
excellent speedups, which is likely due to a mix of prefetching the initial
input files and the I/O patterns of the application. While it was expected that
AFNI would benefit the most from Sea due to its
minimal compute time and the amount of data it generates, it did not benefit
from Sea to the extent that the SPM pipeline did. As per Table~\ref{table:seaneuro-pipelines}, the
AFNI pipeline performs a very high number of glibc calls. While the overhead of an
individual call is likely very minimal, it is possible that the culmination of
all these calls results in a large enough bottleneck that affected both Sea and
Baseline executions to an extensive amount. Moreover, the number of glibc calls
aimed at interacting with data on the Lustre storage are very few. The AFNI pipeline still
obtained an average of 2$\times$ speedup across all experimental conditions.

The FSL Feat pipeline, in contrast, appeared to be the most compute-bound of the three
applications. Not only did it spend an extensive amount of time computing, the
amount of output data generated was least of the three. Due to the
compute-intensive nature of the FSL Feat pipeline, it is expected that it resulted in the
least significant speedups.

The HCP dataset, on average, obtained the greatest speedups with
Sea. Out of the three datasets, HCP has the largest images (see
Table~\ref{table:sea-neuro:data}). Larger individual images mean that they
occupy more page cache space and take longer to flush to Lustre. Unsurprisingly,
the next dataset with the largest images (\SI{282}{\mebi\byte} compressed for a
single image) results in the next largest speedups, followed by PREVENT-AD, the
dataset with the smallest functional images.

For both the pipelines and the datasets, we noticed that obtained speedups
decreased with increased parallelism. We did not modify the default parallelism
level of the applications in our experiments. As a result, each application
process was likely attempting to utilize all cores, resulting in longer wait times
with the increase in parallelism. With this unintentional additional wait time that
is added to the applications due to compute contention, the benefits of using
Sea becomes more limited.

@@

Baseline performance without busy writers was comparable to that of Sea's as can be seen in Figure~\ref{fig:seaneuro:slashbin}. However, with
busy writers, Baseline performance deviated dramatically from the identity
line. We observe that the more Baseline deviates from this identity line, the
greater the speedup. We did occasionally observe slowdowns from using Sea. These slowdowns
may arise from the initial read of the data, as they appeared to occur less
frequently with the SPM pipeline, or due to increased CPU contention caused by Sea's rapid I/O.

\subsection{Overhead of Sea was minimal}

@@

To determine the overhead of Sea, we compared the average performance of Sea with
the performance of pipeline executing entirely within memory
(Figure~\ref{fig:seaneuro:tmpfs}). The results demonstrated that Sea and tmpfs
behave mostly identically. Given these results, we conclude that the
overhead of Sea is minimal and that the differences in makespan may be
due to Sea's initial reading of the data from Lustre.

@@

\begin{figure*}
\begin{subfigure}{0.5\textwidth}
    \centering
    \captionsetup{width=.85\linewidth}
    \includegraphics[width=\columnwidth]{figures/beluga_withflush_nozoom.pdf}%
    \caption{Complete results}\label{fig:seaneuro:belugafullwf}
\end{subfigure}
\begin{subfigure}{0.5\textwidth}
    \centering
    \captionsetup{width=.85\linewidth}
    \includegraphics[width=\linewidth]{figures/beluga_withflush_zoom.pdf}
    \caption{Zoomed between 0 and 2000 seconds}\label{fig:seaneuro:belugazoomwf}
\end{subfigure}
\caption{Makespan comparison between Sea and Baseline on the production cluster with flushing enabled for all files}
\label{fig:seaneuro:beluga-wflush}
\end{figure*}

  We also benchmarked Sea on a production cluster of the Digital Alliance of Canada
  shared with other users,
  using the same datasets and pipelines as for the controlled cluster. 
  Figures~\ref{fig:seaneuro:beluga-noflush} and \ref{fig:seaneuro:beluga-wflush}
  show that although performance in production clusters was highly variable,
  we could still obtain important speedups by using Sea. It is very likely that when we
  executed our experiments without flushing
  (Figure~\ref{fig:seaneuro:beluga-noflush}), Lustre performance was not degraded, 
  resulting in Sea and Baseline performing
  quite similarly to each other.

  Our results with flushing enabled confirmed this theory. When running with
  flushing enabled, which had the additional overhead of having to
  ensure that all the data was copied to Lustre, we noticed that occasionally, we
  obtained very large speedups with both the SPM and AFNI pipelines, suggesting that Lustre
  performance was degraded. The
  maximum speedup observed when running on the production cluster with flushing enabled was
  11$\times$ with the AFNI pipeline processing 1 HCP image. The maximum speedup for the SPM pipeline was
  3.2$\times$, processing 8 ds001545 images.

@@

    Our results show that the speedups obtained from using Sea can be 
    substantial (up to 32$\times$ \TG{this value should be mentioned in the results}) depending on the degradation of the Lustre
    file system. The highest speed-ups observed on a production cluster were
    found to be 11$\times$.

@@

    Although Sea usage did result in performance degradation at times, the magnitude
    of the degradation was less than that of possible speedups
    obtained. Our controlled experiments demonstrated that using Sea with a
    degraded Lustre file system can result in average speedups of up to
    2.5$\times$. Such results indicate that the risk of a slowdown with Sea are
    less important that the potential speedups that can arise.

@@

    Sea's benefit is maximal when it is executed on data-intensive pipelines, that is,
     when the data is written at a rate which exceeds the rate at which the
    compute node's page cache can flush results to Lustre. In our experiments,
    the pipeline which consisted of both the shortest duration and largest
    output size was 16 AFNI pipeline processes preprocessing 16 HCP fMRI images. While the concurrent processes wrote, in total,
    approximately 100~GB of data, the controlled cluster would have had
    approximately 44~GB of dirty cache available per Lustre Object Storage Target (OST) and a maximum amount of
    100~GB of page cache. Since all writes would not have occurred at the same
    time and computation time was extended as a result of CPU contention, even
    our controlled experiments did not showcase the benefits of Sea in a highly
    data-intensive scenario. Since HCP is one of the largest fMRI datasets,
    the only way to augment data intensiveness of the applications is
    through increased parallelism which is limited by the number of available
    cores. As a result, we could only demonstrate the speedups brought by Sea when
    Lustre performance had been degraded.

    Sea cannot speedup applications that are very compute-intensive, as is the case with the FSL Feat pipeline.
    In such cases, the performance bottleneck is the
    CPU time, which Sea does not address. Moreover, depending on the
    application's I/O patterns, it is possible that there will be sufficient
    page cache space and time to flush all the results asynchronously
    to Lustre. However, even with minimal speedups, Sea can still be useful in
    these cases to limit the number of new files created on Lustre, which can be
    particularly useful in instances where users have strict file limits.

@@

    Our results expose the complexities of performance predictions on HPC clusters. We presented results on a dedicated cluster
    where we had
    control over the global cluster usage and were able to demonstrate that
    significant speedups could be obtained when all the Lustre storage devices were
    busy writing other data concurrently. However, even these experiments were
    limited in demonstrating potential speedups with Sea as there were many
    controlled variables, some of which may vary from cluster to cluster. For
    example, the production cluster had a total of 977 nodes and only 38 disks associated with
    their Lustre scratch file system. We can imagine that if 900 of the nodes
    are busy writing to the scratch file system, we would observe a speedup
    larger than what has been reported on our dedicated cluster with only 8 compute
    nodes and 44 disks associated to Lustre. However, these are not the only
    factors that can affect reported results, others can include when the
    experiment were executed, what kind of applications were being executed
    alongside the experiment, scratch file system usage during experiments, and
    what were the application read and write patterns.

@@

    Sea provides prefetching, flushing and eviction mechanisms to further
    improve the speedups that can be obtained. Prefetching allows input data to
    be moved to and manipulated from memory. All of our experiments with the SPM pipeline
    used prefetching as SPM updates the input data via the use of a memory map.
    Without prefetching, updates to the input files would have been performed
    directly on Lustre, thus exhibiting a less important speedup. It was
    favorable to proceed with prefetching, in this case, as the input files
    themselves were not large compared to available memory space. However, in
    cases with very large input files, it may be preferable to maintain the
    memory map to Lustre.

@@

    The ability to use glibc interception over more common alternative methods
    such as kernel-based or FUSE-based file systems was essential in ensuring
    ease-of-use and minimal performance overheads. As both our controlled and
    production environment experiments demonstrated, Sea's overhead is minimal
    and the magnitude of the speedups to be gained by the use of Sea is
    greater than the magnitude of the slowdowns.

@@

    From the perspective of a cluster administrator, Sea is considered useful to
    maintaining the global health of the system. By using available local
    resources, Sea is able to alleviate strain on globally shared cluster
    resources, thus improving the experience for all users. Sea is able to do
    this through the use of eviction which ensures that data that is not
    required never makes it to Lustre.

@@

    To determine the performance gain Sea brings to neuroimaging analyses, we
    must evaluate the value of Sea on a variety of neuroimaging applications.
    For our analysis, we selected different fMRI preprocessing applications as
    fMRI processing has some of the most well-established tools for neuroimaging
    and some of the largest datasets. Of course, different modalities and tools
    may result in vastly different data access patterns and compute times.

@@

    Similarly to datasets, preprocessing pipelines can also vary in duration as
    a result of methodological differences. Thus, we preprocessed each dataset
    using three standard preprocessing pipelines available within the FSL~\cite{fsl},
    AFNI~\cite{cox1996afni}, and SPM~\cite{spm} toolboxes.
    Table~\ref{table:seaneuro-pipelines} provides a reference of the differences
    in computation and data-intensiveness of the different pipelines on a single
    subject for all three of the datasets.

    Each tool was set to only run the functional preprocessing pipeline. For the FSL preprocessing pipeline, 
    we used FEAT configured with many of
    the default parameters. We set the following parameters: slice-timing
    correction (interleaved), intensity normalization and non-linear
    registration to standard space. FSL version 6.0.4 was used.

    %% Add SPM here
    To preprocess the datasets with SPM, we reused the template described in
    \cite{haitas2021age}, with fieldmap correction removed. As the template was
    designed for a specific dataset, we updated the paths to the fMRI and
    anatomical images, as well as the dimensions, such as the number of volumes
    and slices and the TR and TA, for each dataset. Furthermore, we set the
    slice order to interleaved in all cases. SPM version 12 was used.

@@

@@

      \hline
    \end{tabular}}
    \footnotesize{$*$ Measurements taken on the dedicated cluster}
    \caption{Pipeline execution characteristics based on the processing of a single fMRI image using a single application process}
    \label{table:seaneuro-pipelines}
  \end{table*}

    \subsection{Controls}
    Our objective was to measure the speedup brought by Sea on real neuroimaging
    pipelines. Although Sea is agnostic to the internals of the pipelines and
    their output results, incorrect parameter selection and resulting outputs
    may result in I/O patterns that would not be observed had the pipeline been
    executed properly.
